---
title: "Lab-C Replication"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: github_document
---

## Introduction
This notebook replicates the Lab-08 assignment.

## Dependencies
This notebook requires the following packages:

```{r load-packages}
# tidyverse packages
library(dplyr)       # data wrangling tools

# spatial packages
library(nngeo)       # eliminate holes
library(mapview)     # preview spatial data
library(sf)          # spatial data tools

# other packages
library(here)        # file path management
```

## Load Data
This notebook requires two shapefiles, one with precincts in unincorporated parts of St. Louis County and one with precincts in incorporated municipalities:

```{r load-data}
corp <- st_read(here("data", "STLC_ELECT_IncorporatedPrecincts"))
unincorp <- st_read(here("data", "STLC_ELECT_UnincorporatedPrecincts"))
```

## Part 1
### Question 1
First, we need to convert these into a projected coordinate system - I can see in the output of `st_read()` that they are both in the WGS 1984 geographic coordinate system. I'll use Missouri State Plane East:

```{r}
corp <- st_transform(corp, crs = 3602)
unincorp <- st_transform(unincorp, crs = 3602)
```


### Question 2
There are three problems with the unincorporated data - the variable names do not match the incorporated data, and the precinct ID column is character (as opposed to numeric in the incorporated data). There is also an extra column, `WARD`. We'll rename the variables, remove WARD, and then convert the precinct ID column to numeric. The goal is to make the unincorporated data match the incorporated data:

```{r p1-q2}
unincorp %>%
  select(-WARD) %>%
  rename(
    id = PRECINCTID,
    precinct = PRECINCT,
    council = COUNTY_COU,
    muni = MUNI
  ) %>%
  mutate(id = as.numeric(id)) -> unincorp
```

Now we have two matching data frames that can be merged (or bound) together.

### Question 3
To combine these together, we'll use `rbind()`:

```{r p1-q3}
precincts <- rbind(corp, unincorp)
```

We now have a single data frame with all preincts.

### Question 4
Next, we need to visually inspect the `precincts` object to see if it is a geometry collection or if it has polygons. If you look in the global enviornment, you can see that `geometry:sfc_GEOMETRY` appears in the `precincts` object. We'll extract this to `"POLYGON"` before proceeding with the dissolve.

```{r p1-q4}
precincts <- st_collection_extract(precincts, type = "POLYGON")
```

Now we have data prepared for a dissolve.

### Question 5
Next, we'll combine all of the preincts together based on the council district they fall into:

```{r p1-q5-dissolve}
precincts %>%
  select(council) %>%
  group_by(council) %>%
  summarise() -> council
```

We'll preview the changes using `mapview`:

```{r p1-q5-preview}
mapview(council)
```

We can see a variety of holes throughout most of the council districts.

### Question 6
In order to address these holes, we can use `st_remove_holes()` to clean-up the output:

```{r p1-q6}
council <- st_remove_holes(council)
```

We'll preview the changes using `mapview`:

```{r p1-q6-preview}
mapview(council)
```

### Question 7
Finally, we'll create a 1000 meter buffer around the centroid for Council District 3 by:

  * subsetting our observations down to just District 3, then
  * calculating a centroid, then
  * calculating a buffer

```{r p1-q7}
council %>%
  filter(council == 3) %>%
  st_centroid() %>%
  st_buffer(dis = 1000) -> district3
```

We'll preview the changes using `mapview`:

```{r p1-q7-preview}
mapview(district3)
```

